---
title: "Assignment 2--R"
author: "Lyndsey Pohl"
date: ""
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Control flow:  Airport statistics

If your code uses any libraries, load them here.
```{r}
#install.packages("readr")
#I received an error at this line when knitting so I commented it out for that part.
```
```{r}
library(readr)
```
a. Download the data set `airport.csv`.  Read it into R and display the first few lines of data.

```{r}
airportdata = read_csv("airport.csv")
head(airportdata, n=5)


```

This data set contains data on all 135 medium and large airports in the US in 1990.  

(Dataset:  "US Airport Statistics," submitted by Larry Winner, University of Florida.   Dataset obtained from the Journal of Statistics Education (http://www.amstat.org/publications/jse). Accessed 3 June 2015. Used by permission of author.) 

b. **Use control flow** to print a list of airports at which the number of scheduled departures was less than the number of departures performed.

- The point of this assignment is to demonstrate your understanding of control flow, so please use control flow to solve this problem.

```{r}
numRows = dim(airportdata)[1]
for(flight_index in 1:numRows){
  if(airportdata$`Scheduled Departures`[flight_index] < airportdata$`Performed Departures`[flight_index]){
    print(airportdata$Airport[flight_index])
    
  }
  
}
```



c. **Use control flow** to find the average (mean) number of passengers on flights from all the airports in part b.

- Your result should be 1 number:  The total number of passengers from all of the specified airports, divided by the total number of performed departures from all of the specified airports.

- The point of this assignment is to demonstrate your understanding of control flow, so please use control flow to solve this problem.

```{r}
numRows = dim(airportdata)[1]
totpassengers = 0
totdeparts = 0
for(flight_index in 1:numRows){
  if(airportdata$`Scheduled Departures`[flight_index] < airportdata$`Performed Departures`[flight_index]){
#    print(airportdata$Airport[flight_index])
      totdeparts = totdeparts + airportdata$`Performed Departures`[flight_index]
      totpassengers = totpassengers +airportdata$Passengers[flight_index]
    
  }
  
}

meanPassengers = totpassengers / totdeparts
meanPassengers
```

d.  Optional: 	If you know a more efficient way to do parts b and c in R, that's great!  Do that here and use the result to check your work.  (If you don't know a more efficient way, that's fine.  We'll discuss methods of writing efficient code later in the course.)
```{r}
n<-nrow(airportdata)
#n = dim(airportdata)[1]
i=1
count = 0
sum = 0
while (i<=n) {
  if(airportdata[i,3]<airportdata[i,4])
  {sum = sum + airportdata[i,5]
  count = count + airportdata[i,4]
  #print(airportdata[i,1])
  }
  i<-i+1
}

meandeparts = sum/count
meandeparts
```


-	Submit your .Rmd file and knitted .pdf or .docx file (.pdf preferred) to GitHub.
